---
title: "younus_wrangling.Rmd"
author: "Saleem Younus"
date: "9/20/22"
output: 
  html_document: 
    theme: dark
    
---

```{r, warning = FALSE, message = FALSE}
library(tidyverse)  
library(dplyr)
```

## Exercise 1:  Apply DeMorgan’s Law using !

```{r}
txhousing_demorgan <- txhousing %>% 
  filter( (year==2013 | year==2014) )
txhousing_demorgan 
```

## Exercise 2: Distributive Law 

```{r}
txhousing_distributive <- txhousing %>% 
  filter( (sales>=3000 | ( listings>4000 & median>=150000 )))

(txhousing_distributive)

```

## Exercise 3: Percentile Values
```{r}
quantile(txhousing$median, c(.05, .95), na.rm = TRUE)


```

## Exercise 4: Extreme housing prices 2012 - 2014
```{r}
txhousing_price_extremes = txhousing %>%
  filter((median<77525 | median>196700), year >= 2012, year <= 2014)
txhousing_price_extremes =subset (txhousing_price_extremes, select = -inventory)
txhousing_price_extremes
```
Exercise 5: How many houses are sold at what average price?
```{r}
txhousing_augmented <- 
  mutate(
    txhousing,
    percent_sold = (sales) / (listings)*100,
    average_price = (volume) / (sales))

glimpse(txhousing_augmented)
```

## Exercise 6: Total and Average Sales Price by Year
```{r, warning = FALSE, message = FALSE}
sales_by_year = txhousing %>% 
  group_by(year) %>% 
  summarize(avg_sale_price = mean(volume, na.rm=TRUE),
            total_sales = sum(sales, na.rm=TRUE))

sales_by_year
```

## Exercise 7: Number and Value of sales in 2014
```{r}
basic_housing = txhousing %>% 
  mutate(txhousing, average_price = volume/sales) %>%
  filter(year == 2014, city == "Dallas") %>% 
  select(city, year, month, sales, volume, average_price) %>%
  arrange(desc(average_price))
(basic_housing)
```
## Exercise 8: Total Sales in the Top 4 Texas Cities in 2012
```{r}
tophousing <- txhousing %>%
  filter(year == 2010 & city == "Houston" | city == "San Antonio" | city =="Austin" | city == "Dallas") %>%
  group_by(city) %>%
  summarize(total_sales = sum(sales, na.rm = TRUE))

(tophousing)
```

## Exercise 9: Merging previous table with Population 
```{r}
texas_populations <- tibble(
  city = c("Houston" , "San Antonio" , "Austin" , "Dallas"), 
  population = c(2100263 , 1327407 , 1197816 , 790390)
)

texas_big_cities <- tophousing %>% 
  inner_join(texas_populations, by = "city")
texas_big_cities




```

